RNApyro

### 2012-10-03 Recomb/methods_RECOMB.tex
started recomb
added raw equations to recomb

%!TEX root = main_RECOMB.tex
\section{Introduction}
\label{sec:introduction}

>>>>

%!TEX root = main_RECOMB.tex
\section{Methods}
\label{sec:methods}

Given:
\begin{itemize}
	\item an RNA sequence $s$ of length $n$
	\item $s[i]:=$ nucleotide at position $i$ in $s$
	\item $\displaystyle
		S[i]:=\left\{
		\begin{array}{rl}
			-1 & \text{If position }i \text{ unpaired}\\
	 		k  & \text{If position }i \text{ paired with position }k
		\end{array}\right.
		$\\
		S[-1] = -1, special value allowing to compute $S[i-1]$ when $i\equiv 0$
	\item $B:=\{A, U, G, C\}$
	\item $ \Kron_{a,a'}:=\left\{
		\begin{array}{ll}
			1 &\text{If }a \equiv a'\\
			0 &\text{Else}
		\end{array}\right.$
	\item $\Z{i,j}{m}{\alpha,\beta}:=$ Partition function of the subsequence of $s$ from nucleotides 
	$i$ to $j$ with $m$ mutations when nucleotides at positions $i-1=\alpha$ and $j+1=\beta$
	\item $\Y{i,j}{m}{\alpha,\beta}:=$ Partition function of $s$ excluding subsequence  from $i+1$ to 
	$j-1$ with $m$ mutations when nucleotides at positions $i+1=\alpha$ and $j-1=\beta$ 	
	\item Some function $ES_{ab \to a'b'}$ in function of turner starking energy, isostericity and
	  	some $\alpha$. If we choose $\alpha\times\text{StackEnergy} + (1-\alpha)\text{SumIsostericity}$ 
		we 	have the problem that on any invalid base pair (e.g $GG$) the infinite energy given to the 
		stacking 	will remove all information of the isostericity. 
	\item $EI_{ab,a'b'}$ which take into account the isostericity of base pairs not stacked $ab$ when 
	change to $a'b'$.
\end{itemize}

\subsection{Inside}
\subsubsection{Initial conditions}
$$
	\forall i \in (0,\cdots,n-1):\, \Z{i+1,i}{m}{\alpha,\beta}=\left\{
	\begin{array}{ll}
		1 &\text{If } m = 0\\
		0 &\text{Else }
	\end{array}\right.
$$
\subsubsection{Recursion}
$$
	\Z{i,j}{m}{\alpha,\beta}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}  
      \Z{i+1,j}{m-\Kron_{a',s[i]}}{a',\beta} & \text{If }S{[i]}=-1\\
      \displaystyle
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}  
			 e^{\frac{-ES_{\alpha \beta \to a' b'}}{RT}}
			 \Z{i+1,j-1}{m-\Kron_{a'b',s[i]s[j]}}{a',b'}&
			 \text{Elif }S{[i]}=j \land S{[i-1]}=j+1\\
			 \displaystyle
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
      \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
   		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
      \Z{i+1,k-1}{m-\Kron_{a'b',s[i]s[k]}-m'}{a',b'}
      \Z{k+1,j}{m'}{b',\beta} & \text{Elif }S[i]=k \land i < k \leq j\\
      0 &\text{Else}
	\end{array}\right.
$$

\subsection{Outside}	
\subsubsection{Initial conditions}
$$
	\Y{-1,j}{m}{X,X}:=
		\displaystyle
	  \Z{j,n-1}{m}{X,X}
$$
\subsubsection{Recursion}
$$
	\Y{i,j}{m}{\alpha,\beta} = \left\{
  \begin{array}{ll}
		\displaystyle
    \sum_{\substack{a'\in \B,\\ \Kron_{a',s[i]}\le m}}
    \Y{i-1,j}{m- \Kron_{a',s[i]}}{a',\beta} &
    \text{Elif }S[i]=-1 \\
    \displaystyle
    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[j]}\le m}}
		 e^{\frac{-ES_{ a' b' \to \alpha \beta }}{RT}}
    \Y{i-1,j+1}{m- \Kron_{a'b',s[i]s[j]}}{a',b'} &
   	 \text{Elif }S{[i]}=j \land S{[i+1]}=j-1\\
		 \displaystyle
		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[i]s[k]}\le m}}
		 \sum_{m'=0}^{m-\Kron_{a'b',s[i]s[k]}}
  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
		 \Y{i-1,k+1}{m- \Kron_{a'b',s[i]s[k]} - m'}{a',b'}
     \Z{j,k-1}{m'}{\beta,b'} &
		 \text{Elif }S{[i]}=k \geq j\\
		 \displaystyle
		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s[k]s[i]}\le m}}
		 \sum_{m'=0}^{m-\Kron_{a'b',s[k]s[i]}}
  		 e^{\frac{-EI_{s[i]s[k],a'b'}}{RT}}
		 \Y{k-1,j}{m- \Kron_{a'b',s[k]s[i]} - m'}{a',\beta}
     \Z{k+1,i-1}{m'}{a',b'} &
		 \text{Elif }-1 < S{[i]}=k < i\\
		 0 & \text{Else}
  \end{array}\right.
$$
\subsection*{}
We must have (if $S[k] = -1)$:
$$
	\Z{0,n-1}{m}{X,X} \equiv     
	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a};\qquad
	\forall k \in \{0,\cdots,n-1\}
$$

\subsection{Probability of a position being a given nucleotide}
Given $x\in B$,
$$
	\mathbb{P}(s[i] = x\mid s, S,m):=\left\{
	\begin{array}{ll}
		\displaystyle
		\frac{
			\Y{i-1,i+1}{m-\Kron_{x,s[k]}}{x,x}
		}{
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S[i] = -1\\
		\displaystyle
 		\frac{
			\displaystyle
			\sum_{\substack{b\in Bases\\\Kron_{bx,s[k]s[i]\leq m}}}
			\sum_{m'=0}^{m-\Kron_{bx,s[k]s[i]}}
			\Y{k-1,i+1}{m-\Kron_{bx,s[k]s[i]-m'}}{b,x}
			\Z{k+1,i-1}{m'}{b,x}
		}{
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S[i]=k<i\\
		\displaystyle
		\frac{
			\displaystyle
			\sum_{\substack{b\in Bases\\\Kron_{xb,s[i]s[k]\leq m}}}
			\sum_{m'=0}^{m-\Kron_{xb,s[i]s[k]}}
			\Y{i-1,k+1}{m-\Kron_{xb,s[i]s[k]-m'}}{x,b}
			\Z{i+1,k-1}{m'}{x,b}
		}{
			\Z{0,n-1}{m}{X,X}
		}
		&\text{If }S[i]=k>i
	\end{array}\right.
$$
